Followers Are Not Enough: A Question-Oriented Approach to Community Detection in Online Social Networks

### 2014-08-02 paper-for_arxiv/methods.tex
minor edits

This last criterion in particular poses a problem because the majority of community detection algorithms developed so far are built to find partitions of a network and few are aimed at finding overlapping communities \cite{Aicher26062014,BaumesGKMP05,PalEtAl05,ZhaWanZha07,Gre07,PhysRevE.77.016107,Lancichinetti2009,PhysRevE.80.016105,Kovacs2010}. Among these methods, even fewer deal with directed or weighted networks. For example, the work of~\cite{PalEtAl05} on clique percolation can account for both features, but not at the same time. A recent method proposed by~\cite{LancichinettiPlos}, OSLOM (Order Statistics Local Optimization Method), is one of the first methods  able to deal with all of these features simultaneously. Their method relies on a fitness function that measures the statistical significance of clusters with respect to random fluctuations, and attempts to optimize this fitness function across all clusters. 
Following~\cite{LancichinettiPlos}, the significance measure is defined as the probability of finding the cluster in a network without community structure. The null model used is basically the same as the one adopted by Newman and Girvan to define modularity, i.e. it is a model that generates random graphs with a given degree distribution. The authors tested their algorithm on different benchmarks (LFR and Girvan-Newman) and real networks (such as the US air transportation and the word association network), and compared its performance against the best algorithms
currently available (i.e the ones mentioned in the introduction), and found excellent results. Moreover, they showed that OSLOM is also able to recognize the absence, and not simply the presence, of community structure, by testing it on random graphs. This feature of the algorithm plays an important role in the analysis of real data networks, since it is not always the case that a community structure is indeed present and it is therefore useful to be able to detect its absence too. 
Therefore, given its versatility and performance with benchmark network, in this paper we used OSLOM to detect \emph{overlapping} communities present in our \emph{weighted} and \emph{directed} network.
%For this paper we use OSLOM since it allowed us to detect \emph{overlapping} communities present in our \emph{weighted} and \emph{directed} network, and its authors showed that it performs very well on various types of artificial benchmark graphs and also on several real networks.

>>>>

This last criterion in particular poses a problem because the majority of community detection algorithms developed so far are built to find partitions of a network and few are aimed at finding overlapping communities \cite{Aicher26062014,BaumesGKMP05,PalEtAl05,ZhaWanZha07,Gre07,PhysRevE.77.016107,Lancichinetti2009,PhysRevE.80.016105,Kovacs2010}. Among these methods, even fewer deal with directed or weighted networks. For example, the work of~\cite{PalEtAl05} on clique percolation can account for both features, but not at the same time. A recent method proposed by~\cite{LancichinettiPlos}, OSLOM (Order Statistics Local Optimization Method), is one of the first methods  able to deal with all of these features simultaneously. Their method relies on a fitness function that measures the statistical significance of clusters with respect to random fluctuations, and attempts to optimize this fitness function across all clusters. 
Following~\cite{LancichinettiPlos}, the significance measure is defined as the probability of finding the cluster in a network without community structure. The null model used is basically the same as the one adopted by Newman and Girvan to define modularity, i.e. it is a model that generates random graphs with a given degree distribution. The authors tested their algorithm on different benchmarks (LFR and Girvan-Newman) and real networks (such as the US air transportation and the word association network), and compared its performance against the best algorithms
currently available (i.e the ones mentioned in the introduction), and found excellent results. Moreover, they showed that OSLOM is also able to recognize the absence, and not simply the presence, of community structure, by testing it on random graphs. This feature of the algorithm plays an important role in the analysis of real data networks, since it is not always the case that a community structure is indeed present and it is therefore useful to be able to detect its absence too. 
Therefore, given its versatility and performance with benchmark network, in this paper we used OSLOM to detect \emph{overlapping} communities present in our \emph{weighted} and \emph{directed} networks.
%For this paper we use OSLOM since it allowed us to detect \emph{overlapping} communities present in our \emph{weighted} and \emph{directed} network, and its authors showed that it performs very well on various types of artificial benchmark graphs and also on several real networks.

### 2014-08-02 paper-for_arxiv/paper.tex
minor edits

\section{Funding}
Elisa Omodei is partially supported by a PhD grant from the Région Ile-de-France (DIM 2011) and by the French National Agency of Research (ANR) through the grant Algopol (ANR-12-CORD-0018).
We are grateful to DARPA for the support of David Darmon.
\section{Acknowledgments}

>>>>

\section{Funding}
%Elisa Omodei is partially supported by a PhD grant from the R\'{e}gion \^{I}le-de-France (DIM 2011) and by the French National Agency of Research (ANR) through the grant Algopol (ANR-12-CORD-0018).
This work was supported by the R\'{e}gion \^{I}le-de-France [DIM 2011 to E.O.]; and by the French National Agency of Research [ANR-12-CORD-0018 to E.O.].
We are grateful to DARPA for the support of David Darmon.

\section{Acknowledgments}

### 2014-08-02 paper-for_arxiv/results.tex
minor edits

As an aside, it is interesting to examine who the most influential users are in these communities, as determined by transfer entropy. Recall that influence in this context refers to a reduction in uncertainty about a follower(s) activity on Twitter given the tweet history of a followee. Said differently, higher transfer entropy, or a higher reduction in uncertainty, implies more influence. Interestingly, among the most influential users in these communities, those with highest transfer entropy on average, we find \@Colorado, which is the state official Twitter account, \@ConnectColorado, a page created to connect Coloradans, and the CBS Denver account, a popular news agency. This means that these accounts had high influence in terms of their followers' activity on Twitter. This is not surprising as these communities discuss Colorado and these accounts provide information on this topic. However, this does provide a check that this measure of activity based influence makes sense and suggests this could be a nice measure for influence on social media if used with other measures. This is discussed more in future work.

Lastly, it is interesting to note that two of the Forbes ``Top 10 Social Media Influencers", % ranked using the total outgoing strength in the activity-based network, 
 happen to be in our network, \emph{viz.}, Ann Tran and Jessica Northey, and transfer entropy also quantified these two users as highly influential.% members, as categorized by transfer entropy. 
This suggests that activity-based or information-theoretic influence may be useful in finding influential members purely based on their temporal tweet history while completely ignoring both tweet content and social status.

>>>>

As an aside, it is interesting to examine who the most influential users are in these communities, as determined by transfer entropy. Recall that influence in this context refers to a reduction in uncertainty about a follower(s) activity on Twitter given the tweet history of a followee. Said differently, higher transfer entropy, or a higher reduction in uncertainty, implies more influence. Interestingly, among the most influential users in these communities, those with highest transfer entropy on average, we find \@Colorado, which is the state official Twitter account, \@ConnectColorado, a page created to connect Coloradans, and the CBS Denver account, a popular news agency. This means that these accounts had high influence in terms of their followers' activity on Twitter. This is not surprising as these communities discuss Colorado and these accounts provide information on this topic. However, this does provide a check that this measure of activity-based influence makes sense and suggests this could be a useful measure for influence on social media if used with other measures. This is discussed more in future work.

Lastly, it is interesting to note that two of the Forbes ``Top 10 Social Media Influencers", % ranked using the total outgoing strength in the activity-based network, 
 happen to be in our network, \emph{viz.}, Ann Tran and Jessica Northey, and transfer entropy also quantified these two users as highly influential.
% members, as categorized by transfer entropy. 
This suggests that activity-based or information-theoretic influence may be useful in finding influential members purely based on their temporal tweet history while completely ignoring both tweet content and social status.
